Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2024-01-10 E-Hypergraphs-LICS24.tex
typoes sec 5 and 6

CHRIS: REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ MORPHISMS, AND WHY!
Because extended cospans have a more general notion of interface,  we need to make some adjustments to DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics}.
The first thing is that we need to account for is the existence of \textit{internal} interfaces,  and in particular that internal interfaces are not necessarily preserved during rewriting.
The former means that rewrite rules become of the form
\[
\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
\] where we use the fact that cospans 
\[
    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
\]
and
\[
    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j
\]
encode the same data and we will further omit the arrows $0 \to 0 \to \mathcal{L}$.
That is, a rewrite rule is a pair of cospans of e-hypergraphs with matching \textit{external} interfaces and arbitrary \textit{internal} interfaces.
To illustrate the non-preservation of internal interfaces, consider the rewrite rule above and the following e-hypergraph with interfaces to perform the rewrite on
% \[\langle x \xrightarrow{} x' \xrightarrow{} \mathcal{L} \xleftarrow{} y' \xleftarrow{} y, x \xrightarrow{} x'' \xrightarrow{} \mathcal{R} \xleftarrow{} y'' \xleftarrow{} y \rangle \]
% such that $x' \not \cong x''$ and $y' \not \cong y''$ and consider the following e-hypergraph to perform the rewrite on
\[
    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k
\] such that there is a matching $m : \mathcal{L} \to \mathcal{G}$  and assume $i' \not \cong i''$ and $j' \not \cong j''$.
When removing the occurence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of $\mathcal{G}$ should be modified to exclude the vertices that were a part of (the image of) $\mathcal{L}$'s strict internal interfaces.
Then, when gluing $\mathcal{R}$ into the created context, we need to introduce strict internal interfaces of $\mathcal{R}$ into this context.
In particular, it means that the internal interfaces of the resulting cospan of e-hypergraphs
\[
    \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k 
\]
not necessarily coincide with the internal interfaces of the initial cospan.
% The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
% Note that the outermost interfaces are preserved.

@@

To account for such changes in the interfaces we require additional structure on DPOI rewriting for $\catname{EHyp({\Sigma})}$ which is summarised in the definition below.
\begin{definition}[Extended DPOI (EDPOI) rewriting]
\label{def:dpoi-e}
Given a span of morphisms 
\[
    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
\]
in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites to} $\mathcal{H}$ \textit{with external interface} $n + k$ \textit{via a rewrite rule} 
\[
    \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
\]
if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram such that the two marked squares are pushouts.
% \[\begin{tikzcd}
% 	&&&& {i + j} \\
% 	{l_i+l_o} && L &&&& R && {r_i+r_o} \\
% 	\\
% 	{g_i+g_o} && {G^{\urcorner}} && C && {^{\ulcorner}H} && {h_i+h_o} \\
% 	\\
% 	&&&& {n+m}
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-3]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-7]
% 	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=2-3, to=4-3]
% 	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=4-5]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=2-7, to=4-7]
% 	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-5]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-7]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-3]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-3]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-7]
% 	\arrow[from=2-1, to=2-3]
% 	\arrow["{[f_i,f_o]}", from=2-1, to=4-1]
% 	\arrow[from=4-1, to=4-3]
% 	\arrow[from=2-9, to=2-7]
% 	\arrow[from=4-9, to=4-7]
% 	\arrow["{[h_i,h_o]}"{description}, from=2-9, to=4-9]
% 	\arrow[from=6-5, to=4-9]
% 	\arrow[from=6-5, to=4-1]
% 	\arrow["{[i_l,j_l]}"{description}, from=1-5, to=2-1]
% 	\arrow["{[i_r,j_r]}"{description}, from=1-5, to=2-9]
% \end{tikzcd}\]

% https://q.uiver.app/#q=WzAsMTEsWzIsMCwiTCJdLFs0LDAsInhfe2V4dH0gKyB5X3tleHR9Il0sWzIsMiwiR157XFx1cmNvcm5lcn0iXSxbNCwyLCJDIl0sWzQsNCwibl97ZXh0fSttX3tleHR9Il0sWzAsMCwibF9pK2xfbyJdLFswLDIsImdfaStnX28iXSxbNiwwLCJSIl0sWzYsMiwiXntcXHVsY29ybmVyfUgiXSxbOCwyLCJoX2kraF9vIl0sWzgsMCwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw4LCIiLDAseyJjb2xvdXIiOlswLDYwLDYwXX1dLFszLDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzcsOCwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbMSw3LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxMCw3XSxbOSw4XV0=
\[
 \scalebox{0.75}{
    \tikzfig{combinatorial_semantics/DPOI_square}
 }
\]
\end{definition}
To guarantee that the rewriting yields a monogamous directed acyclic e-hypergraph we also need to introduce a revised version of a boundary complement.

@@

For monogamous extended cospans 
\[
    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
\] 
and 
\[
    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
\] and mono $f : \mathcal{L} \to \mathcal{G}$, a pushout complement 
\[
    i + j \to \mathcal{L}^{\bot} \to G^{\urcorner}
\] 
as depicted in the square below
\[
\scalebox{0.75}{

@@

is a \textit{boundary} complement if $[c_1,c_2]$ is mono 
 and there exist $d_1 : n \to C$ and $d_2 : k \to C$ 
 making the above triangle commute and such that
there exists a \textit{not necessarily well-typed} mda-cospan

@@

where $l$ and $r$ are
\[
    l = n' \setminus (i' \setminus i)
\]
and
\[
    r = k' \setminus (j' \setminus j)
\]
Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed apart from the part that has a pre-image in $i + j$ there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
\[
    D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
\]
Similarly for $r \to \mathcal{L}^{\bot}$ which we denote as $g_2$.
Then there is a morphism $n \to l$ because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
Similarly for $k \to r$.
\end{definition}

@@

We are now ready to define the \textit{convex} rewriting for e-hypergraphs with interfaces which we will later use to show the equivalence between $\textsf{PROP}(\Sigma)$ and $\WellTypedMdaEcospans$.
It will be analogous to the one from Definition~\ref{def:convex_dpo} but again, to make sure the resulting cospan of e-hypergraph is monogamous we will have to construct internal interfaces separately.
We will also limit matches to convex down-closed ones both to guarantee an mda result and so that the result is a well-defined e-hypergraph.

\begin{definition}[Convex EDPOI rewriting]
    Let $\mathfrak{R}$ be a set of EDPOI rewrite rules.
    Then, given 
    \[
        \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k
    \] 
    and 
    \[
        \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
    \] in $\catname{EHyp(\Sigma)}$ with the external interfaces $n + k$, $\mathcal{G}$ \textit{rewrites convexly into} $\mathcal{H}$ \textit{with the outermost interface} $n + k$ --- notation $( \mathcal{G} \xleftarrow{} n + k) \Rrightarrow_{\mathfrak{R}}^{+} (\mathcal{H} \xleftarrow{} n + k)$ --- if there exists a rule 
    \[
        \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
    \] 
    with the external interfaces $i + j$ in $\mathfrak{R}$ and object $\mathcal{C}$ and cospan arrows $i + j \xrightarrow{} \mathcal{C} \xleftarrow{} n + k$ 
    such that the diagram of Definition~\ref{def:dpoi-e} commutes and its marked squares are pushouts, and the following conditions hold

    \begin{enumerate}
        \label{dpoi-e:assumptions}
        \item $m : \mathcal{L} \to \mathcal{G}$ is a convex down-closed match;
        \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement in the leftmost pushout.
    \end{enumerate}
\end{definition}

% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
%  \[
%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
% \]
\begin{align*}
&n'' = (i'' \setminus i) + l\\
&k'' = (j'' \setminus j) + r\\
&n'' \to \mathcal{H} = [f_1,f_2] \text{ where }\\
&f_1 : i'' \setminus i \to \mathcal{R} \to \mathcal{H}\\ 
&f_2 : l \to \mathcal{C} \to \mathcal{H} 
\end{align*}
Since both $f_1$ and $f_2$ are monos, their copairing is also a mono. 
Similarly for $k'' \to \mathcal{H}$.

@@

The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}. 
\begin{proposition}[Weak normal form]
For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
\[
	f = f_1 + \ldots + f_n
\]
 such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
\end{proposition}

>>>>

CHRIS: REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ MORPHISMS, AND WHY!
Because extended cospans have a more general notion of interface, DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
First, the existence of \textit{internal} interfaces needs to be accounted for,  and in particular that internal interfaces are not necessarily preserved during rewriting.
The former means that rewrite rules become of the form
\[
\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
\] where we use the fact that cospans 
\[
    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
\]
and
\[
    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j
\]
encode the same data and we will further omit the arrows $0 \to 0 \to \mathcal{L}$.
That is, a rewrite rule is a pair of cospans of e-hypergraphs with matching \textit{external} interfaces and arbitrary \textit{internal} interfaces.
To illustrate the non-preservation of internal interfaces, consider the rewrite rule above and the following e-hypergraph with interfaces to perform the rewrite on
% \[\langle x \xrightarrow{} x' \xrightarrow{} \mathcal{L} \xleftarrow{} y' \xleftarrow{} y, x \xrightarrow{} x'' \xrightarrow{} \mathcal{R} \xleftarrow{} y'' \xleftarrow{} y \rangle \]
% such that $x' \not \cong x''$ and $y' \not \cong y''$ and consider the following e-hypergraph to perform the rewrite on
\[
    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k
\] such that there is a matching $m : \mathcal{L} \to \mathcal{G}$  and assume $i' \not \cong i''$ and $j' \not \cong j''$.
When removing the occurence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of $\mathcal{G}$ should be modified to exclude the vertices that were a part of (the image of) $\mathcal{L}$'s strict internal interfaces.
Then, when gluing $\mathcal{R}$ into the created context, we need to introduce strict internal interfaces of $\mathcal{R}$ into this context.
In particular, it means that the internal interfaces of the resulting cospan of e-hypergraphs
\[
    \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k 
\]
not necessarily coincides with the internal interfaces of the initial cospan.
% The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
% Note that the outermost interfaces are preserved.

@@

To account for such changes in the interfaces, consider additional structure on DPOI rewriting for $\catname{EHyp({\Sigma})}$, summarised in the definition below.
\begin{definition}[Extended DPOI (EDPOI) rewriting]
\label{def:dpoi-e}
Given a span of morphisms 
\[
    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
\]
in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites to} $\mathcal{H}$ \textit{with external interface} $n + k$ \textit{via a rewrite rule} 
\[
    \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
\]
if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram such that the two marked squares are pushouts.
% \[\begin{tikzcd}
% 	&&&& {i + j} \\
% 	{l_i+l_o} && L &&&& R && {r_i+r_o} \\
% 	\\
% 	{g_i+g_o} && {G^{\urcorner}} && C && {^{\ulcorner}H} && {h_i+h_o} \\
% 	\\
% 	&&&& {n+m}
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-3]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-7]
% 	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=2-3, to=4-3]
% 	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=4-5]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=2-7, to=4-7]
% 	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-5]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-7]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-3]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-3]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-7]
% 	\arrow[from=2-1, to=2-3]
% 	\arrow["{[f_i,f_o]}", from=2-1, to=4-1]
% 	\arrow[from=4-1, to=4-3]
% 	\arrow[from=2-9, to=2-7]
% 	\arrow[from=4-9, to=4-7]
% 	\arrow["{[h_i,h_o]}"{description}, from=2-9, to=4-9]
% 	\arrow[from=6-5, to=4-9]
% 	\arrow[from=6-5, to=4-1]
% 	\arrow["{[i_l,j_l]}"{description}, from=1-5, to=2-1]
% 	\arrow["{[i_r,j_r]}"{description}, from=1-5, to=2-9]
% \end{tikzcd}\]

% https://q.uiver.app/#q=WzAsMTEsWzIsMCwiTCJdLFs0LDAsInhfe2V4dH0gKyB5X3tleHR9Il0sWzIsMiwiR157XFx1cmNvcm5lcn0iXSxbNCwyLCJDIl0sWzQsNCwibl97ZXh0fSttX3tleHR9Il0sWzAsMCwibF9pK2xfbyJdLFswLDIsImdfaStnX28iXSxbNiwwLCJSIl0sWzYsMiwiXntcXHVsY29ybmVyfUgiXSxbOCwyLCJoX2kraF9vIl0sWzgsMCwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw4LCIiLDAseyJjb2xvdXIiOlswLDYwLDYwXX1dLFszLDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzcsOCwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbMSw3LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxMCw3XSxbOSw4XV0=
\[
 \scalebox{0.75}{
    \tikzfig{combinatorial_semantics/DPOI_square}
 }
\]
\end{definition}
To guarantee that the rewriting yields a monogamous directed acyclic e-hypergraph boundary complement also requires a revised definition.

@@

For monogamous extended cospans 
\[
    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
\quad\text{ and }\quad
    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
\] and mono $f : \mathcal{L} \to \mathcal{G}$, a pushout complement 
\[
    i + j \to \mathcal{L}^{\bot} \to G^{\urcorner}
\] 
as depicted in the square below
\[
\scalebox{0.75}{

@@

is a \textit{boundary} complement if $[c_1,c_2]$ is mono 
 and there exist $d_1 : n \to C$ and $d_2 : k \to C$ 
 making the above triangle commute and such that
 DAN: THIS (AND SUCH THAT) IS EXTREMELY UNCLEAR, PLEASE REWORD
there exists a \textit{not necessarily well-typed} mda-cospan

@@

where $l$ and $r$ are
\[
    l = n' \setminus (i' \setminus i)
\quad\text{ and }\quad
    r = k' \setminus (j' \setminus j).
\]
Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
\[
    D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
\]
Similarly for $r \to \mathcal{L}^{\bot}$, which we denote as $g_2$.
Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
Similarly for $k \to r$.
\end{definition}

@@

We are now ready to define the \textit{convex} rewriting for e-hypergraphs with interfaces which we will later use to show the equivalence between $\textsf{PROP}(\Sigma)$ and $\WellTypedMdaEcospans$.
It will be analogous to the one from Definition~\ref{def:convex_dpo} but again, to make sure the resulting cospan of e-hypergraphs is monogamous we will have to construct internal interfaces separately.
We will also limit matches to convex down-closed ones both to guarantee an mda result and so that the result is a well-defined e-hypergraph.

\begin{definition}[Convex EDPOI rewriting]
    Let $\mathfrak{R}$ be a set of EDPOI rewrite rules.
    Then, given 
    \[
        \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k
    \quad 
    \text{and} 
    \quad
        \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
    \] in $\catname{EHyp(\Sigma)}$ with the external interfaces $n + k$, $\mathcal{G}$ \textit{rewrites convexly into} $\mathcal{H}$ \textit{with the outermost interface} $n + k$ --- notation $( \mathcal{G} \xleftarrow{} n + k) \Rrightarrow_{\mathfrak{R}}^{+} (\mathcal{H} \xleftarrow{} n + k)$ --- if there exists a rule 
    \[
        \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
    \] 
    with the external interfaces $i + j$ in $\mathfrak{R}$ and object $\mathcal{C}$ and cospan arrows $i + j \xrightarrow{} \mathcal{C} \xleftarrow{} n + k$ 
    such that the diagram of Definition~\ref{def:dpoi-e} commutes and its marked squares are pushouts, and the following conditions hold

    \begin{enumerate}
        \label{dpoi-e:assumptions}
        \item $m : \mathcal{L} \to \mathcal{G}$ is a convex down-closed match;
        \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement in the leftmost pushout.
    \end{enumerate}
\end{definition}
and
% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
%  \[
%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
% \]
\begin{align*}
&n'' = (i'' \setminus i) + l\\
&k'' = (j'' \setminus j) + r\\
&n'' \to \mathcal{H} = [f_1,f_2] \text{ where }\\
&f_1 : i'' \setminus i \to \mathcal{R} \to \mathcal{H}\\ 
&f_2 : l \to \mathcal{C} \to \mathcal{H} 
\end{align*}
Since both $f_1$ and $f_2$ are monos, their copairing is also a mono. 
Similarly for $k'' \to \mathcal{H}$.

@@

The equations of Figure~\ref{fig:string-equations}---excepting the commutativity equation---can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}. 
\begin{proposition}[Weak normal form]
For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
\[
	f = f_1 + \ldots + f_n
\]
 such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
\end{proposition}
